\section{Discussion}
\label{sec:discussion}

\subsection{Reconciling the Conflicting Reports}

The three-axis decomposition lets us reconcile the conflicting
reports in the hybrid AR/DDLM literature.  Papers that report ``AR
planning helps DDLM'' likely measured the planner-content-trust
axis with a well-matched planner-and-adapter capacity --- i.e.\ a
configuration where the upstream planner produces content the DDLM
trusts at the prevailing capacity setting.  Papers that report
``AR planning hurts DDLM'' likely measured the format-brittleness
axis without controlling for prefix shape: prepending a plan-shaped
scaffold of \emph{any} content ($66\text{--}68\%$ at base, versus
$74\%$ for no prefix) is enough to land in the
``hurts'' regime.  Papers that report ``diffusion sampling
diversity is preserved'' likely measured the right thing under a
recipe that lands at v2-style capacity; papers that report
``diversity collapses'' likely had a different recipe that pushed
past the capacity sweet spot we identify in
Table~\ref{tab:axis1_v2_v3}.

Without the three-axis frame, these are five conflicting claims.
Within the frame, they are five projections of a multi-axis
failure surface, each consistent with the others when the omitted
axes are held fixed.

\subsection{Implication: Distillation of Inference-Time Mechanisms}

The Track 2 design-iteration finding has a broader implication:
weight-space distillation of inference-time mechanisms is
design-sensitive in a way that produces clean negative results
that masquerade as architectural impossibilities.

Distilling a technique that aggregates over multiple stochastic
samples (\cmaj{}) into a single deterministic forward pass
requires the surgery to match the temporal structure of the
original mechanism.  For semi-AR diffusion, ``temporal structure''
means \emph{which sub-block fires the modification}.  Getting this
wrong (last-block-only when the answer is upstream of the
modification) produces \ctwoc{} $= 70.5\%$ across a $3.25\times$
capacity bump --- a clean negative whose surface reading is ``the
adapter cannot do this''.  Getting it right (commit on sub-blocks
$2$--$4$ of $4$) recovers \ctwoc{} $= 79.0\%$, within sampling
error of the target, on the same data.  The architecture did not
change; the temporal placement did.

This is a general lesson for parameter-efficient distillation of
sampling-based mechanisms: the placement of the LoRA in the
sampler's schedule is a first-class design variable, and a
plateau across capacity is not by itself evidence that the
distillation is impossible.

\subsection{Future Work}

\paragraph{Cross-task transfer.}  Do these axes generalize beyond
GSM8K~\citep{cobbe2021gsm8k}?  A targeted ablation on MATH-Easy
and ARC-Challenge would test whether the three-axis decomposition
holds on other reasoning benchmarks.  The diversity-expansion
finding in particular is suspicious-looking on a single benchmark
and deserves a multi-task replication.

\paragraph{Reranker baseline.}  Train a small classifier over the
$5$ \cmaj{} branches as a \emph{compute-time} alternative to
v3's \emph{param-time} commit.  If a reranker matches \ctwoc{}
$= 79\%$, weight-space distillation is not uniquely useful and
the simpler compute-time recipe is preferable.  If the reranker
falls short, the param-time mechanism is doing something
distinct.

\paragraph{Diversity-mechanism test.}  Does base \LLaDA{} at
$t{=}0.9$ reproduce the post-Track-1 $5/5$-same rate of $47.5\%$?
If yes, Track 1 acts as an implicit temperature regularizer and
the diversity-expansion finding reduces to that.  If no, the
training-time surface randomization has produced real
content-level diversity that temperature does not match.  One
short eval; deferred to revision.

\paragraph{Content-trust adapter.}  The planner-content-trust
axis (\S\ref{sec:axis2}) is characterized but not fixed.  A
content-trust LoRA --- one that varies plan content quality at
fixed surface shape, complementing our prefix-robust dataset
which varies surface shape at fixed content --- is a natural
follow-up, especially with \Qwen{1.5B} and \Qwen{7B} planners
sweeping the upstream-content axis explicitly.

\paragraph{Adaptive mode router (Paper 2).}  The
structural-separation finding motivates a learned router that
switches between mode-specific LoRAs based on input.  The
capacity-tradeoff curve (v2 vs.\ v3) gives the router a real
decision space: given an input prompt, which adapter should fire?
We sketched the design in \code{sfumato/PLAN.md} as the E1 track
and leave it to a follow-up.

\subsection{Limitations}

We close with what this paper does not do.  All evaluations are
single-seed at $N{=}200$, so our headline numbers carry binomial
sampling noise of $\pm 3$--$5$\,pp; multi-seed variance is not
characterized in this draft.  We did not produce a mechanistic
LCH-centroid figure for the
prefix-robust LoRA, which would harden the structural-separation
thesis from behavioural to mechanistic; the JVP feasibility spike
failed because PEFT wraps LoRA in
\code{lora\_magnitude\_vector} ModuleDict and we deferred to a
follow-up.  All experiments use a single base diffusion model
(\LLaDA{}-8B-Instruct~\citep{nie2025llada}) and a single
benchmark (GSM8K-test~\citep{cobbe2021gsm8k}); the
generalization claims in the abstract are constrained
accordingly.  These exposures and a few smaller ones are listed in
\appref{app:open}.
