\section{Introduction}
\label{sec:introduction}

Hybrid pipelines that pair an autoregressive (AR) language model with
a discrete-diffusion language model (DDLM) for chain-of-thought
reasoning have produced contradictory empirical reports.  Some
authors find that an AR planner improves DDLM
denoising~\citep{ye2024diffusionofthought,zhao2025d1};
others find that the same pattern degrades it~\citep{huang2025dcolt};
and reports on whether DDLM sampling diversity is preserved or
collapses under fine-tuning likewise disagree~\citep{liu2025ccdd}.
The picture is muddied further by the fact that recent
DDLMs~\citep{nie2025llada,gong2024diffullama,arriola2025bd3lms} share
a tokenizer family and a semi-AR sampler but are evaluated on
different benchmarks and at different prefix conditions, so the
underlying disagreement is hard to isolate.

We started this project from a narrow observation in our own prior
diagnostic work (E4): on \LLaDA{}-8B-Instruct~\citep{nie2025llada},
prepending a literal \code{"Plan: "} string to a GSM8K
question~\citep{cobbe2021gsm8k} costs $8$\,pp of accuracy versus the
no-prefix baseline ($74\% \rightarrow 66\%$), even though the prefix
contains no content at all.  The damage is format-shape, not plan
quality.  This single number was inconsistent with the simplest
``AR plan helps DDLM'' story and is also inconsistent with the
simplest ``AR plan hurts DDLM'' story: the cost is real, but the
mechanism is not the planner.  It is the prefix surface itself.

This paper argues that the conflicts in the literature collapse once
the failure surface is decomposed into orthogonal axes that can be
measured and intervened on independently.  We identify three.

\begin{enumerate}
    \item \textbf{Interface-format brittleness}: how much the DDLM's
          accuracy drops when the question is wrapped in any
          plan-shaped scaffold, content-free or otherwise.
    \item \textbf{Planner-content trust}: how much the DDLM uses the
          actual content of an upstream AR plan, conditional on the
          prefix shape being absorbed.
    \item \textbf{Sampling-diversity preservation}: whether
          fine-tuning the DDLM collapses, preserves, or expands the
          variety of branches produced by stochastic sampling, on
          which compute-time consensus mechanisms such as
          self-consistency~\citep{wang2023selfconsistency} depend.
\end{enumerate}

On GSM8K-test (N=200) with \LLaDA{}-8B-Instruct and
\Qwen{\{0.5B,1.5B\}}-Instruct~\citep{qwen2024report} planners, we
characterize each axis with a small ($r{=}8$)
LoRA~\citep{hu2022lora,mangrulkar2022peft} and report the following.

First, interface-format brittleness is trainably fixable.  A
prefix-robust LoRA on \LLaDA{} with full FFN module coverage (v3,
22M trainable parameters) lifts no-prefix \ctwo{} from $74\% \to
73.0\%$ (within sampling error) while pulling the empty-prefix
\code{"Plan: "} condition from $66\% \to 74\%$, flattening the
no-prefix-vs-empty-prefix gap from $8$\,pp to within $1$\,pp.  A
partial-coverage variant (v2, 4/7 modules, 10M parameters) achieves
slightly weaker single-shot but stronger branch-vote performance,
exposing a no-prefix-vs-\cmaj{} capacity tradeoff.

Second, planner-content trust is capacity-dependent in opposite
directions across planners.  At v2 capacity, a $1.5$B AR planner
helps ($60\% \to 67\%$, $+7$\,pp) and a $0.5$B planner still hurts
($64\% \to 60\%$).  At v3 capacity this inverts: the $0.5$B
planner improves ($64\% \to 65\%$, $+5$\,pp) and the $1.5$B
planner catastrophically regresses ($60\% \to 54\%$, $-13$\,pp).
Capacity differentially modulates how much the model trusts
upstream content, and the optimal capacity depends on the planner.

Third, format-augmented training expands sampling diversity rather
than collapsing it.  After the prefix-robust LoRA (v2) plus a
late-block commit adapter, the rate at which all $5$ stochastic
branches agree drops from $51.5\%$ to $47.5\%$ on test, and mean
unique answers per problem rises from $1.83$ to $2.07$ ($+13\%$).
\cmaj{} accuracy holds at $81.5\%$ on dev, and \cmajc{} reaches
$82.0\%$ (commit v2) / $82.5\%$ (commit v3) on test.  Diversity
goes up, accuracy is preserved or improved.

Fourth, weight-space distillation of the consensus mechanism
(\cmaj{} into a single forward pass) is design-sensitive in a way
that is easy to miss.  A late-block, answer-span commit adapter
fails to bridge the $6$\,pp gap between greedy \ctwo{} and \cmaj{},
plateauing at \ctwoc{}$=70.5\%$ across two iterations and a
$3.25\times$ capacity bump.  A multi-block, full-response variant
recovers \ctwoc{}$=79.0\%$, within sampling error of the $80\%$
pre-registered target.  Disentangling ablations attribute $+4$\,pp
of the lift to where commit fires in the diffusion schedule and
$+2$\,pp to full-response supervision; full-response training on its
own (with last-block-only commit) adds nothing.

\clearpage
\paragraph{Contributions.}
\begin{enumerate}
    \item \textbf{A three-axis decomposition of hybrid AR/DDLM
          reasoning failure.}  Interface-format brittleness,
          planner-content trust, and sampling-diversity preservation
          are independently measurable on GSM8K with \LLaDA{} and
          \Qwen{\{0.5B,1.5B\}}.  The decomposition reconciles
          conflicting reports in the literature
          (\S\ref{sec:axis1}, \S\ref{sec:axis2}, \S\ref{sec:axis3},
          \S\ref{sec:distill}, \S\ref{sec:discussion}).
    \item \textbf{Interface-format brittleness is trainably fixable
          with a small ($r{=}8$) prefix-robustness LoRA, with a
          measurable capacity tradeoff.}  A v2 ($4/7$ FFN modules,
          $10$M parameters) and a v3 ($7/7$ FFN modules, $22$M
          parameters) variant trade $+2.5$\,pp on no-prefix \ctwo{}
          for $-2.0$\,pp on \cmaj{} branch-vote, locating a
          single-shot-vs-ensemble Pareto frontier
          (\S\ref{sec:axis1}).
    \item \textbf{Planner-content trust is capacity-dependent in
          opposite directions across planners --- a
          previously-unmeasured axis.}  A $0.5$B planner improves
          $+5$\,pp at v3 capacity while a $1.5$B planner regresses
          $-13$\,pp, with the directions reversed at v2 capacity
          (\S\ref{sec:axis2}).
    \item \textbf{Format-augmented LoRA training expands sampling
          diversity rather than collapsing it.}  $5/5$-branch
          agreement drops $51.5\% \to 47.5\%$ and mean unique
          answers per problem rises $1.83 \to 2.07$, while \cmaj{}
          accuracy is preserved.  We connect this to encoder-collapse
          literature as the inverse phenomenon
          (\S\ref{sec:axis3}).
    \item \textbf{Consensus distillation into a DDLM is
          design-sensitive, not architecture-limited.}  Late-block
          answer-span surgery plateaus at \ctwoc{}$=70.5\%$ across
          two iterations; multi-block ($n{=}3$) full-response
          surgery recovers \ctwoc{}$=79.0\%$.  Disentangling
          ablations attribute most of the lift ($+4$\,pp) to block
          coverage; full-response loss alone (with last-block-only
          commit) adds $+0.0$\,pp.  The two design errors mask each
          other (\S\ref{sec:distill}).
    \item \textbf{Param-time and compute-time consensus compose.}
          \cmajc{} (cmaj with commit-LoRA per branch) reaches
          $82.5\%$ on test for the v3 commit and $82.0\%$ for the
          v2 commit, $+3.5$\,pp and $+3$\,pp respectively over
          base test \cmaj{} of $79.0\%$.  Pre-registered
          prediction \#3 (no double-dip) is violated upward in
          both cases.  A \Qwen{0.5B} self-consistency
          baseline~\citep{wang2023selfconsistency} on the same eval
          reaches only $40.5\%$, ruling out generic self-consistency
          as the explanation for \LLaDA{}'s diffusion \cmaj{}
          advantage (\S\ref{sec:distill}, \S\ref{sec:rebuttal}).
\end{enumerate}

\paragraph{Framing: decomposition, not unification.}
This paper does not propose a single architecture or training recipe.
It argues that the empirical record on hybrid AR/DDLM reasoning is
muddied because the failure surface is multi-dimensional and prior
work has measured projections of it without controlling for the
other axes.  The contribution is the decomposition itself, plus
trainable interventions on two of the three axes that demonstrate
the axes are real and individually addressable.  The third axis
(planner-content trust) is characterized but not fixed; we leave a
content-trust adapter to future work.

\paragraph{Status.}
All numbers are measured on GSM8K-test, $N{=}200$, fixed seed $0$
unless otherwise noted, with \LLaDA{}-8B-Instruct at $k{=}64$
denoising steps and a fixed regex answer extractor that tolerates
both \code{Answer:} and \code{\#\#\#\#} formats.  All four LoRA
adapters (Track 1 v2, Track 1 v3, Track 2 v2, Track 2 v3) and the
three derived datasets are public on the Hugging Face Hub.  Total
compute spend on a single RTX~4090 spot pod is approximately
$\$3.50$ (\appref{app:compute}).  \appref{app:hyperparameters}
lists exact training configurations; \appref{app:prereg} reproduces
the pre-registration scorecard, including a prediction violated in
the unexpected (good) direction; \appref{app:open} flags open
exposures (multi-seed variance, \cmajc{} with the v3 commit, and a
mechanistic LCH centroid analysis) that we did not close before
this draft.
