\section{Reviewer-Resilience: Is This Just Self-Consistency?}
\label{sec:rebuttal}

The most natural alternative explanation for everything reported
above is generic AR self-consistency~\citep{wang2023selfconsistency}:
sample $b$ branches at temperature $t$, take the majority vote, and
get the same accuracy lift the diffusion \cmaj{} mechanism appears
to deliver.  If true, the diffusion-specific story --- that
\LLaDA{}'s within-block uncertainty produces genuinely different
reasoning paths rather than surface jitter --- would be unnecessary.

We rule this out empirically with a matched baseline.

\subsection{Baseline}

\Qwen{0.5B}-Instruct~\citep{qwen2024report} self-consistency at
$b{=}5$, $t{=}0.7$, on the same GSM8K-test eval ($N{=}200$, fixed
problem indices, fixed seed):

\begin{table}[h]
\centering
\caption{AR self-consistency baseline against \LLaDA{} \cmaj{} on
the same eval ($N{=}200$, GSM8K-test, $b{=}5$, $t{=}0.7$).  Generic
self-consistency does not explain the diffusion \cmaj{} advantage.}
\label{tab:rebuttal_sc}
\smallskip
\begin{tabular}{lr}
\toprule
Setup & Accuracy \\
\midrule
\Qwen{0.5B} self-consistency, $b{=}5$, $t{=}0.7$ & $40.5\%$ \\
\LLaDA{} \cmaj{} on test (base, no LoRA) & $79.0\%$ \\
\LLaDA{} \cmaj{} (Track 1 v2 + commit v2) & $\mathbf{82.0\%}$ \\
\bottomrule
\end{tabular}
\end{table}

The gap between \Qwen{0.5B} self-consistency and \LLaDA{}
\cmaj{} is $38.5\text{--}41.5$\,pp.  Generic self-consistency at
matched $b$, $t$, and eval does not explain the advantage.

\subsection{Reading}

The \Qwen{0.5B} planner is the same model used for the
\cthree{}\code{p Q-0.5B} condition elsewhere in this paper.  At
that scale, AR self-consistency does not produce enough genuinely
different reasoning paths to vote effectively: most of the
$5$ branches differ in surface phrasing but not in the answer.
\LLaDA{}'s diffusion sampler resolves within-block uncertainty
jointly across many tokens, producing trajectories that diverge
in their \emph{reasoning} rather than only in their phrasing.
This is consistent with the diversity-expansion finding in
\S\ref{sec:axis3}: the diffusion sampler at $t{=}0.7$ is already
exploring enough trajectory variety that \cmaj{} has real
disagreement to aggregate over, and Track 1 training expands the
distribution further rather than collapsing it.

This rebuttal is narrow.  It does not show that diffusion
\cmaj{} beats AR self-consistency at all scales --- a $7$B
\Qwen{} planner with self-consistency would likely close some of
the gap.  It shows that at the matched-base-model scale relevant
to this paper, the simplest alternative explanation is empirically
inadequate.
