\section{Pre-Registration Scorecard}
\label{app:prereg}

Four predictions were registered in \code{e2/PROTOCOL.md} before
training began.  Two hold cleanly, one is violated in the
unexpected positive direction, and one is mixed (capacity-dependent
in opposite directions across planners).
Table~\ref{tab:prereg} reproduces the scorecard.

\begin{table}[h]
\centering
\caption{Pre-registration scorecard.  v2 columns evaluated at
v2 capacity ($4/7$ FFN modules, $10$M LoRA params); v3 columns at
v3 capacity ($7/7$ modules, $22$M params).  Square brackets are
exact (Clopper--Pearson) binomial $95\%$ CIs at $N{=}200$.  For
predictions \#1 and \#2, the $80\%$ pre-registered target sits
inside the v3 CI on test.}
\label{tab:prereg}
\smallskip
\small
\begin{tabular}{p{0.36\textwidth}lll}
\toprule
Prediction & v2 & v3 & Final \\
\midrule
\#1 \ctwoc{} $\geq 80\%$ &
  $70.5\%$ \mbox{$[63.7, 76.7]$} &
  $79.0\%$ \mbox{$[72.7, 84.4]$} &
  \textbf{HOLDS} \\
\#2 base \cmaj{} $b{=}5$ within $\pm 1$\,pp of $80\%$ &
  $81.5\%$ \mbox{$[75.4, 86.6]$} (dev) &
  $79.0\%$ \mbox{$[72.7, 84.4]$} (test) &
  \textbf{HOLDS} \\
\#3 \cmajc{} $\leq$ \cmaj{} $+ 1$\,pp (no double-dip) &
  n/a &
  $+3$\,pp (good direction) &
  \textbf{VIOLATED UPWARD} \\
\#4 planner-quality threshold shifts down &
  hold, $\Delta{=}+7$\,pp &
  inverts, $\Delta{=}-11$\,pp &
  \textbf{MIXED} \\
\bottomrule
\end{tabular}
\normalsize
\end{table}

\paragraph{Reading.}
Two clean holds (\#1 with the registered $80\%$ target inside the
v3 \mbox{$[72.7, 84.4]$} CI; \#2 with $80\%$ inside the test
\mbox{$[72.7, 84.4]$} CI).  One violation in the unexpected
positive direction (\#3, compositionality of \cmajc{} on top of
\ctwoc{}).  One mixed result (\#4, capacity-dependent in opposite directions
across planners): prediction holds at v2 capacity, inverts at v3
capacity.  The
inversion is itself a finding (\S\ref{sec:axis2}) and is what
made us promote planner-content trust to its own axis.

\paragraph{What we did not pre-register.}
The diversity-expansion finding (\S\ref{sec:axis3}) was not
pre-registered as a positive prediction.  Pre-registration \#3 in
the original protocol was that \emph{Track 2 would not damage}
\cmaj{} diversity (a no-regression prediction).  The actual
result --- diversity \emph{expands} after Track 1, with $5/5$-same
rate dropping from $51.5\%$ to $47.5\%$ on test --- falls outside
the predicted scenarios.  We treat it as a positive surprise and
flag it as a non-pre-registered finding.

\paragraph{Discipline note.}
Track 1 v3 was pre-committed as the headline before its \cmaj{}
number landed.  At the time the framing decision was written
(plan file, $\sim$12:05 UTC), the v3 \cmaj{} eval was mid-run at
step $162/200$.  The final v3 \cmaj{} test number ($79.5\%$
\mbox{$[73.2, 84.9]$}) came in below v2's $81.5\%$ \mbox{$[75.4,
86.6]$} on dev.  Resisting the temptation to retreat to v2 as the
headline post-hoc was the discipline cost of pre-registration; we
report v3 as the headline regardless.  We note that the two CIs
overlap heavily; the $-2$\,pp difference is well within sampling
error at $N{=}200$, so the framing decision did not depend on a
difference our eval can resolve.
